%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\topmargin=-0.5in % Moves the top of the document 1 inch above the default
\textheight=8.75in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=0in % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

%\address{Rutgers University and University of Michigan}

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professors Galin Jones and Faming Liang \\
    Co-Editors, {\em Journal of Computational and Graphical Statistics}
         }
%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

%\begin{center}
%\large
%\end{center}
%\vfill

\signature{Walter Dempsey\\
University of Michigan \\
Department of Biostatistics\\
M4057 SPH II \\
1415 Washington Heights \\
wdem@umich.edu} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\date\today

\opening{Dear Drs. Jones and Liang:}

Kindly find enclosed the revised version of the manuscript ID JCGS-22-236, "Recurrent event analysis in the presence of real-time high frequency data via random subsampling."   We have revised the article to address the referees’ requests.  The reports called for a major revision to include related work on modeling event data with functional data predictors as well as historical functional linear models, include empirical evidence and associated discussion of the computational benefits of our proposal, as well as clarify several technical details in the theoretical sections and provide important intuition/sketches of the proofs. The new version complies with these requests.

We believe the work presents an important advance that gives researchers a simple tool to analyze recurrent event data in the presence of high-frequency sensors data.  Attached please find a detailed response to the referees outlining how the changes to the manuscript address the concerns raised in their reports.

Thank you for your consideration. We look forward to hearing back from you.

\closing{Sincerely,}


\end{letter}

\end{document}


